---
title: "Exercise 8"
author: "Nikolaus Czernin"
output: pdf_document
fig_width: 6 
fig_height: 6
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
library("knitr")
# library("ROCit")
# library("ISLR")
# library("klaR")
# library("glmnet")
# install.packages("gclus")
library("gclus")
library("tidyverse")

set.seed(11721138)
```

## Loading, preprocessing & splittting
```{r}
data(ozone)

# perform train set split
train_idx <- sample(1:nrow(ozone), nrow(ozone)%/%3*2, replace=FALSE)
train <- ozone[train_idx, ]
test <-  ozone[-train_idx, ]

# ozone
```

```{r}
getsplines <- function(x, nknots=2, M=4){
  # nknots ... number of knots -> placed at regular quantiles
  # M ... M-1 is the degree of the polynomial
  n <- length(x)
  # X will not get an intercept column
  X <- matrix(NA, nrow=n, ncol=(M - 1) + nknots)
  for (i in 1:(M - 1)){ 
    X[,i] <- x^i 
    }
  # now the basis functions for the constraints:
  quant <- seq(0, 1, 1 / (nknots + 1))[c(2 : (nknots + 1))]
  qu <- quantile(x,quant)
  for (i in M : (M + nknots - 1)){
    X[,i] <- ifelse(x - qu[i - M + 1] < 0, 0, (x - qu[i - M + 1])^(M - 1))
  }
  list(x=x, X=X, quantiles=quant, xquantiles=qu)
}

```


# Task 1: Creating and plotting splines
```{r}
plotspl <- function(splobj, ..., title=""){
  vertical.lines <- splobj$xquantiles %>% unname()
  splobj %>%  
    .$X %>% 
    data.frame() %>% 
    mutate(x=X1) %>% 
    pivot_longer(-c(x), values_to = "hx") %>% 
    ggplot(aes(x=x, y=hx)) +
      geom_line() +
      theme_minimal() +
      facet_wrap(~name, scales="free") +
      geom_vline(xintercept=vertical.lines, 
                 colour="skyblue", 
                 linetype = "longdash") +
      ggtitle(title) 
  }

# getsplines(ozone$Temp) %>% 
#   plotspl(main="Temp")
```


```{r}
# These lines create splines for all columns, not just temp
# splines <- train %>% 
#   colnames() %>% 
#   lapply(function(name){
#     list(name=name, splines=getsplines(train[[name]]))
#   })
# 
# splines %>% 
#   lapply(function(spline) plotspl(spline$splines, title=spline$name))

# Creating splines for the temperature
splines.temp.train <- getsplines(train$Temp)
splines.temp.test <- getsplines(test$Temp)
plotspl(splines.temp.train, "Temperature Splines")
# plotspl(splines.temp.test, "Temperature Splines")

```

# Task 2: Fitting a linear model

```{r}
# fit the model
training.data <- data.frame(splines.temp.train$X, Ozone=train$Ozone)
model1 <- lm(Ozone ~ ., data=training.data)
summary(model1)
# make predictions
train.idx <- train$Temp %>% sort(index.return=TRUE) %>% .$ix
test.idx <- test$Temp %>% sort(index.return=TRUE) %>% .$ix
yhat.train <- predict(model1, training.data)

plot(train$Temp[train.idx], yhat.train[train.idx], ylab="Predictions", main="Predicted ozone concentration per temperature", col="blue", type="l")
```

# Task 3: Plotting predictions vs observations

```{r}


plot(train$Temp[train.idx], yhat.train[train.idx], ylab="Predictions", main="Predicted ozone concentration per temperature", col="blue", type="l")

test.data <- data.frame(splines.temp.test$X, Ozone=test$Ozone)
yhat.test <- predict(model1, test.data)
lines(test$Temp[test.idx], yhat.test[test.idx], col="green")

```

# Task 4: Extending the data
```{r}
plot(train$Temp[train.idx], yhat.train[train.idx], ylab="Predictions", main="Predicted ozone concentration per temperature", col="blue", type="l",
     xlim=c(0, 120), 
     ylim=c(-200, 30)
     )

lines(test$Temp[test.idx], yhat.test[test.idx], col="green")


newtemp <- seq(0, 120)

newtemp %>% 
  getsplines() %>% 
  .$X %>% 
  as.data.frame() %>% 
  rename_all(str_replace_all, "V", "X") %>% 
  predict(model1, .) %>% 
  lines(newtemp, ., col="orange")
```

# Task 5: Knots at custom points
```{r}
getsplines <- function(x, nknots=2, M=4, knots=NULL){
  # nknots ... number of knots -> placed at regular quantiles
  # M ... M-1 is the degree of the polynomial
  n <- length(x)
  # X will not get an intercept column
  X <- matrix(NA, nrow=n, ncol=(M - 1) + nknots)
  for (i in 1:(M - 1)){ 
    X[,i] <- x^i 
    }
  # now the basis functions for the constraints:
  if (is.null(knots)){
    print("a")
    # create knots from quantiles
    quant <- seq(0, 1, 1 / (nknots + 1))[c(2 : (nknots + 1))]
    qu <- quantile(x,quant)
  } else {
    # if custom knots have been given, use them as knots
    qu = knots
    # and calculate the quantile percentages of the knots
    quant <- ecdf(x)(knots)
  }
  for (i in M : (M + nknots - 1)){
    X[,i] <- ifelse(x - qu[i - M + 1] < 0, 0, (x - qu[i - M + 1])^(M - 1))
  }
  list(x=x, X=X, quantiles=quant, xquantiles=qu)
}

```

The function is now modiffied to accept custom knot values as an optional 
argument. 

```{r}
plot(train$Temp[train.idx], yhat.train[train.idx], ylab="Predictions", main="Predicted ozone concentration per temperature", col="blue", type="l",
     xlim=c(0, 120),
     ylim=c(-40, 40)
     )


# get the knots of the train data splines
training_knots <- splines.temp.train$xquantiles %>% unname()

newtemp <- seq(0, 120)
newtemp %>% 
  getsplines(knots = training_knots) %>% 
  .$X %>% 
  as.data.frame() %>% 
  rename_all(str_replace_all, "V", "X") %>% 
  predict(model1, .) %>% 
  lines(newtemp, ., col="orange")

lines(train$Temp[train.idx], yhat.train[train.idx], col="blue")
lines(test$Temp[test.idx], yhat.test[test.idx], col="green")

```


# Task 6: Repeating the analyses with the log response
```{r}

training.data <- data.frame(splines.temp.train$X, Ozone=train$Ozone)
# if specify using log for every variable in the data
# i also add a small constant to every cell to avoid lm(0)
model2 <- lm(Ozone ~ log(X1) + log(X2) + log(X3) + log(X4) + log(X5),
             data=training.data+1e-12)
# predict train data
yhat.train <- predict(model2, training.data) %>% exp()
# predict test data
test.data <- data.frame(splines.temp.test$X, Ozone=test$Ozone)
yhat.test <- predict(model2, test.data+1e-12) %>% exp()


plot(train$Temp[train.idx], yhat.train[train.idx], ylab="Predictions", main="Predicted ozone concentration per temperature", col="blue", type="l",
     xlim=c(0, 120),
     ylim=c(0, 10e10)
     )


newtemp <- seq(0, 120)
newtemp %>% 
  getsplines(knots = training_knots) %>% 
  .$X %>% 
  as.data.frame() %>% 
  rename_all(str_replace_all, "V", "X") %>% 
  predict(model2, .) %>% 
  exp() %>% 
  lines(newtemp, ., col="orange")

lines(train$Temp[train.idx], yhat.train[train.idx], col="blue")
lines(test$Temp[test.idx], yhat.test[test.idx], col="green")

```

Now the predicted ozone concentrations are all above 0, but they explode 
for temperatures higher than 80. 






